---
title: "HW_5"
author: "Austin Pesina"
date: "3/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ISLR)
library(MASS)
library(boot)
```

## Problem 3  

We now review k-fold cross validation.  
  
**(a) Explain how $k$-fold cross validation is implemented.**  
  
$k$-fold cross validation takes the amount of observations, $n$, and splits them into $k$ non-overlapping groups. The length of the groups is approximately $\frac{n} {k}$. These "folds"are the validation set, while everything else becomes the training set. The test error is achieved by averaging the $k$ errors.
  
**(b) What are the advantages and disadvantages of $k$-fold cross validation relative to:** 

 **i. The validation set approach?  **  
 
The validation set approach's estimate of the test error rate can be a high variable because the observations included in the training/validation set are randomly split. Because not all of the observations are used to fit the model, the error rate may overestimate the test error for the model on the entire data set.     

 **ii. LOOCV?**  
   
LOOCV has less bias. The model is fit to training data using $n-1$ observations. Since most, but not all, of the data is used, a less variable MSE is produced. While the validation set approach can produce different outcomes based on randomness in the splitting process, LOOCV always yields the same results because splits are based on 1 observation each time. The biggest disadvantage is LOOCV is computationally intensive because the model is fit $n$ times.

## Problem 5

In Chapter 4 we used logistic regression to predict the probability of `default` using `income` and `balance` on the [Default](https://rdrr.io/cran/ISLR/man/Default.html) data set. We will now estimate the test error of this logistic regression model using the validation set approach. Do not forget to set a random seed before beginning your analysis.

**(a) Fit a logistic regression mode that uses `income` and `balance` to predict `default`.**  
  
```{r}
attach(Default)
set.seed(1)
fit.glm <- glm(default ~ income + balance, data=Default, family = "binomial")
summary(fit.glm)
```
  
**(b) Using the validation set approach, estimate the test error of this model. In order to do this, you must perform the following steps:**  
**i. Split the sample set into a training set and a validation set.**

```{r}
train <- sample(dim(Default)[1], dim(Default)[1]/2)
```

**ii. Fit a multiple logistic regression model using only the training observations.**

```{r}
fit.glm <- glm(default ~ income + balance, data=Default, family = "binomial", subset = train)
summary(fit.glm)
```

**iii. Obtain a prediction of default status for each individual set by computing the posterior probability of `default` for that individual, and classifying the individual to the default category if the posterior probability is greater than 0.5.**

```{r}
probs <- predict(fit.glm, newdata = Default[-train, ],type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
```

**iv. Compute the validation set error, which is the fraction of the observations in the validation set that are misclassified.**

```{r}
mean(pred.glm != Default[-train, ]$default)
```

**(c) Repeat the process in (b) three times, using three different splits of the observations into a training set and a validation set. Comment on the results obtained.**

```{r}
train <- sample(dim(Default)[1], dim(Default)[1]/2)
fit.glm <- glm(default ~ income + balance, data=Default, family = "binomial", subset = train)
probs <- predict(fit.glm, newdata = Default[-train, ],type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)

```
```{r}


train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```

```{r}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```

The different split all produce different results that vary based on the random splitting of observations in the training and validation sets.  

**(d) Now consider a logistic regression model that predicts the probability of `default` using `income`, `balance`, and a dummy variable for `student.` Estimate the test error for this model using the validation set approach. Comment on whether or not including a dummy variable for `student` leads to a reduction in the test error rate.**

```{r}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance + student, data = Default, family = "binomial", subset = train)
pred.glm <- rep("No", length(probs))
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```

The dummy variable, `student`, did not have a great effect on the test error rate. 

  
## Problem 6

**We continue to consider the use of a logistic regression model to predict the probability of `default` using `income` and `balance` on the `Default` data set. In particular, we will now compute estimates for the standard errors of the `income` and `balance` logistic regression coefficients in two different way: (1) using the bootstrap, and (2) using the standard formula for computing the standard errors in the `glm()` function. Do not forget to set a random seed before beginning your analysis.**


**(a) Using the `summary()` and `glm()` function, determine thew estimated standard errors for the coefficients associated with `income` and `balance` in a multiple logistic regression model that uses both predictors.**

```{r}
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial")
summary(fit.glm)
```
The `glm()` estimates of standard errors are $\beta_0 = 0.4348$, $\beta_1 = 4.985*10^{-6}$, and $\beta_2 = 2.274*10^{-4}$.
**(b) Write a function, `boot.fn()`, that takes as input the `Default` data set as well as an index of the observations, and that outputs the coefficient estimates for `income` and `balance` in the multiple logistic regression model.**

```{r}
boot.fn <- function(data, index){
  fit <- glm(default ~income +balance, data = data, family = "binomial", subset=index)
  return(coef(fit))
}
```

**(c) Use the `boot()` function together with your `boot.fn()` function to estimate the standard errors of the logistic regression coefficients for `income` and `balance.`**

```{r}
boot(Default, boot.fn, 1000)
```
The bootstrap estimates of standard errors are $\beta_0 = 0.4365$, $\beta_1 = 4.8621*10^{-6}$, and $\beta_2 = 2.3091*10^{-4}$.

**(d) Comment on the estimated standard errors obtained using the `glm()` function and using your bootstrap function.**  

The bootstrap and glm() standard errors are both very similar.  

##Problem 9

We will now consider the [Boston](https://www.cs.toronto.edu/~delve/data/boston/bostonDetail.html) housing data set from the `MASS` library.  

**(a)Based on this data set, provide an estimate for the population mean of medv. Call this estimate $\hat{\mu}$.**

```{r}
attach(Boston)
mu.hat <- mean(medv)
mu.hat
```

**(b) Provide an estimate of the standard error of $\hat{\mu}$. Interpret this result.**

```{r}
se.hat <- sd(medv)/sqrt(dim(Boston)[1])
se.hat
```

**(c) Now estimate the standard error of $\hat{\mu}$ using the bootstrap. How does this compare to your answer from (b)?**

```{r}
boot.fn <- function(data, index){
  mu <- mean(data[index])
  return (mu)
}
boot(medv, boot.fn, 1000)
```
The bootstrap estimated standard error of $\hat{\mu}$ is 0.4137, which is close to the estimate of 0.4088 found in part b.  

**(d) Based on your bootstrap estimate from (c), provide a 95% confidence interval for the mean of `medv.` Compare it to the results obtained using `t.test(Boston$medv)`.** *Hint: You can approximate a 95% confidence interval using the formula* $[\hat{\mu}-2SE(\hat{\mu}), \hat{\mu}+2SE(\hat{\mu})]$

```{r}
t.test(medv)
```
```{r}
ci.mu.hat <- c(22.5348-2*0.4137, 22.5348+2*0.4137)
ci.mu.hat
```
The bootstrap confidence interval is close to the t.test() function.  

**(e) Based on this data set, provide an estimate, $\hat{\mu}_{medv}$, for the median value of `medv` in the population.**

```{r}
median.medv <- median(medv)
median.medv
```

**(f) We now would like to estimate the standard error of $\hat{\mu}_{medv}$. Unfortunately, there is no simple formula for computing the standard error of the median. Instead, estimate the standard error of the median using the bootstrap. Comment on your findings.**

```{r}
boot.fn <- function(data, index){
  mu <- median(data[index])
  return(mu)
}
boot(medv, boot.fn, 1000)
```
The estimated median value via bootstrap is 21.2, the same as what we got with the `median()` function. The standard error is 0.3737 which is relatively small.

**(g) Based on this data set, provide an estimate for the tenth percentile of `medv` in Boston suburbs. Call this quantity $\hat{\mu}_{0.1}$**

```{r}
percentile.hat <- quantile(medv, c(0.1))
percentile.hat
```

**(h) Use the bootstrap to estimate the standard error of $\hat{\mu}_{0.1}$. Comment on your findings.**

```{r}
boot.fn <- function(data, index){
  mu <- quantile(data[index],c(0.1))
  return(mu)
}
boot(medv, boot.fn, 1000)
```
We get an estimated 10th percentile of 12.75, the same value as in part g. The standard error is also small, at 0.4854.