---
title: "Ch 4 Homework"
author: "Austin Pesina"
date: "3/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r getdata}
library(ISLR)
library(MASS)
library(class)
library(corrplot)
```

## Problem 10

This question should be answered using the [Weekly](https://rdrr.io/cran/ISLR/man/Weekly.html) data set, which is part of the [ISLR] package. This data is similar in nature to the `Smarket` data from this chapter's lab, except that it contains 1,089 weekly returns for 21 years, from the beginning of 1990 to the end of 2010.  
  
```{r Weekly, message=FALSE, warning=FALSE}
attach(Weekly)
```
  
**(a) Produce Some numerical and graphical summaries of the `Weekly` data. Do there appear to be any patterns?**  
```{r 10-a}
names(Weekly)
pairs(Weekly)
cor(Weekly[,-9])
corrplot(cor(Weekly[,-9]), method = "square")
```
  
There is a correlation between Volume and Year.  
  
  
**(b) Use the full data set to perform a logistic regression with `Directions` as the response and the five lag variables plus `Volume` as predictors. Use the summary function to print the results. Do any of the predictors appear to be statistically significant? If so, which ones?** 

```{r 10-b}
glm.fit <- glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume, data=Weekly, family = binomial)
summary(glm.fit)
```
Lag2 is statistically significant with a p-value of $< 0.05$. 
  
  
**(c) Compute the confusion matrix and overall fraction of correct predictions. Explain what the confusion matrix is telling you about the types of mistakes made by logistic regression.**

```{r 10-c}
glm.probs <- predict(glm.fit, type='response')
glm.probs[1:10]
#glm.probs
contrasts(Direction)

glm.pred <- rep("Down", 1089)
glm.pred[glm.probs > 0.5]="Up"
table(glm.pred, Direction)
```
  
There were 54 observations of `Down` predicted correctly and 557 observations of `Up` predicted correctly. This gives us a 56.11% accuracy.  $\frac{54+557} {54+48+430+557} = 0.5611$ 
  
  
**(d) Now fit the logistic regression model using a training data period from 1990 to 2008, with `Lag2` as the only predictor. Compute the confusion matrix and the overall fraction of correct predictions for the held out data (that is, the data from 2009 and 2010).**

```{r 10-d}
training <- (Year <= 2008)
Weekly.08 <- Weekly[!training,]
Weekly.fit <- glm(Direction ~ Lag2, data = Weekly, subset = training, family = "binomial")
weekly.probs <- predict(Weekly.fit, Weekly.08, type = "response")
weekly.pred <- rep("Down", length(weekly.probs))
weekly.pred[weekly.probs > 0.5]="Up"
Direction.08 <- Direction[!training]
table(weekly.pred, Direction.08)
mean(weekly.pred == Direction.08)
```
  
From 1990 to 2008, the weekly change was predicted with 62.5% accuracy with this model.

**(e) Repeat (d) using LDA.**

```{r 10-e}
lda.fit <- lda(Direction ~ Lag2, data = Weekly, subset = training, family = "binomial")
lda.pred <- predict(lda.fit, Weekly.08)
lda.class <- lda.pred$class
table(lda.class, Direction.08)
mean(lda.class==Direction.08)
```
  
Using LDA, we also get a success rate of 62.5%.  

**(f) Repeat (d) using QDA.**

```{r 10-f}
qda.fit <- qda(Direction ~ Lag2, data = Weekly, subset = training)
qda.fit
qda.class <- predict(qda.fit, Weekly.08)$class
table (qda.class, Direction.08)
mean(qda.class==Direction.08)
```
  
With QDA, our accuracy drops down to 58.65%.  
  
  
**(g) Repeat (d) using KNN with K = 1.**

```{r 10-g}
weekly.training <- as.matrix(Lag2[training])
weekly.test <- as.matrix(Lag2[!training])
train.Direction <- Direction[training]


set.seed(1)
knn.pred <- knn(weekly.training, weekly.test, train.Direction, k=1)
table(knn.pred, Direction.08)
mean(knn.pred == Direction.08)
```
  
Using KNN where $K = 1$, we get an accuracy of 50%.  

**(h) Which of these methods appears to provide the best results on this data?**  
  
The best methods are Logistic Regression and Linear Discriminant Analysis; both have an accuracy of 62.5%  

**(i) Experiment with different combinations of predictors, including possible transformations and interactions, for each of the methods. Report the variables, method, and associated confusion matrix that appears to provide the best results on the held out data. Note that you should also experiment with values for K in the KNN classifier.**  
  
### QDA
```{r 10-i-QDA}
qda.fit2 <- qda(Direction ~ Lag1 + Lag2 + Lag5, data = Weekly, subset = training)
qda.fit2
qda.class2 <- predict(qda.fit2, Weekly.08)$class
table (qda.class2, Direction.08)
mean(qda.class2==Direction.08)
```
  
### LDA
```{r 10-i-LDA}
lda.fit2 <- lda(Direction ~ Lag1 + Lag2 + Lag5, data = Weekly, subset = training, family = "binomial")
lda.pred2 <- predict(lda.fit2, Weekly.08)
lda.class2 <- lda.pred2$class
table(lda.class2, Direction.08)
mean(lda.class2==Direction.08)
```
  
### 1NN
```{r 10-i-1NN}
weekly.training2 <- cbind(Lag1, Lag2, Lag5)[training,]
weekly.test2 <- cbind(Lag1, Lag2, Lag5)[!training,]
train.Direction2 <- Direction[training]
set.seed(1)

knn.pred1 <- knn(weekly.training2, weekly.test2, train.Direction2, k=10)
table(knn.pred1, Direction.08)
mean(knn.pred1 == Direction.08)
```

### 10NN
```{r 10-i-10NN}
knn.pred2 <- knn(weekly.training2, weekly.test2, train.Direction2, k=10)
table(knn.pred2, Direction.08)
mean(knn.pred2 == Direction.08)
```
  
### 100NN
```{r 10-i-100NN}

knn.pred3 <- knn(weekly.training2, weekly.test2, train.Direction2, k=100)
table(knn.pred3, Direction.08)
mean(knn.pred3 == Direction.08)
```
  
It seems like LDA and KNN, where $K = 1$ are our best fits. Note that as K increases, our accuracy decreases.  


## Problem 11

In this problem, you will develop a model to predict whether a given car gets high or low gas mileage based on the [Auto](https://rdrr.io/cran/ISLR/man/Auto.html) data set.  

```{r}
attach(Auto)
```

**(a) Create a binary variable, `mpg01`, that contains a 1 if `mpg` contains a value above its median, and a 0 if `mpg` contains a value below its median. You can compute the median using the `median()` function. Note you may find it helpful to use the `data.frame()` function to create a single data set containing both `mpg01` and the other `Auto` variables.**  

```{r 11-a}
mpg01 <- rep(0, length(Auto$mpg))
mpg01[Auto$mpg > median(Auto$mpg)] <- 1
Auto <-  data.frame(Auto, mpg01)

```
  
  
**(b) Explore the data graphically in order to investigate the association between `mpg01` and the other features. Which of the other features seem most likely to be useful in predicting `mpg01`? Scatterplots and boxplots may be useful tools to answer this question. Describe your findings.**  

```{r 11-b}
corrplot(cor(Auto[,-9]), method = "square")
```
`Cylinders`, `Displacement`, and `Weight` have the strongest association with mpg01.There is also a strong association with `year`, as newer cars have better fuel efficiency.
  
  
**(c) Split the data into a training set and a test set.**

```{r 11-c}
set.seed(1234)
num_train <- nrow(Auto) * 0.7
in_Train <- sample(nrow(Auto), size = num_train)
Auto_training <- Auto[in_Train, ]
Auto_testing <- Auto[-in_Train, ]
```

**(d) Perform LDA on the training data in order to predict `mpg01` using the variables that seemed most associated with `mpg01` in (b). What is the test error of the model obtained?**  
  
```{r 11-d}
Auto.lda <- lda(mpg01 ~ displacement + weight + cylinders + year, data = Auto_training)
Auto.prediction <- predict(Auto.lda, Auto_testing)
table(Auto.prediction$class, Auto_testing$mpg01)
mean(Auto.prediction$class == Auto_testing$mpg01)
```
This model has an error rate of 13.6%.    
  
  
**(e) Perform QDA on the training data in order to predict `mpg01` using the variables that seemed most associated with `mpg01` in (b). What is the test error of the model obtained?**  

```{r 11-e}
Auto.qda <- qda(mpg01 ~ displacement + weight + cylinders + year, data = Auto_training)
test.prediction <- predict(Auto.qda, Auto_testing)
table(test.prediction$class, Auto_testing$mpg01)
mean(test.prediction$class == Auto_testing$mpg01)
```
  
The QDA model has an accuracy of 85.6%, or an error rate of 14.4%.  
  
  
**(f) Perform logistic regression on the training data in order to predict `mpg01` using only the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?**

```{r 11-f}
Auto.logreg <- glm(mpg01 ~ displacement + weight + cylinders + year, data = Auto_training, family = "binomial")
predicted.probs <- predict(Auto.logreg, Auto_testing, type = "response")
preds <- rep(0, dim(Auto_testing)[1])
preds[predicted.probs > 0.5] = 1
table(preds, Auto_testing$mpg01)
mean(preds == Auto_testing$mpg01)
```
   
Logistic Regression has an accuracy of 88.1% or an error rate of 11.9%.  

**(g) Perform KNN on the training data, with several values of K, in order to predict mpg01. Use only the variables that seemed most associated with `mpg01` in (b). What test errors do you obtain? Which value of K seems to perform the best on this data set?**  

```{r 11-g}
train.x2 <- cbind(Auto_training$displacement, Auto_training$weight, Auto_training$cylinders, Auto_training$year)
test.x2 <- cbind(Auto_testing$displacement, Auto_testing$weight, Auto_testing$cylinders, Auto_testing$year)
train.y2 <- cbind(Auto_training$mpg01)
```
   
### For K = 1:
```{r 11-g-1NN}
knn.auto1 <- knn(train.x2, test.x2, train.y2, k=1 )
table(knn.auto1, Auto_testing$mpg01)
mean(knn.auto1 == Auto_testing$mpg01)
```
The accuracy of this model is 84.7%, or an error rate of 15.3%.
  
### For K = 10
```{r 11-g-10NN}
knn.auto10 <- knn(train.x2, test.x2, train.y2, k=10)
table(knn.auto10, Auto_testing$mpg01)
mean(knn.auto10 == Auto_testing$mpg01)
```
The accuracy of $K = 10$ is 88.1% or an error rate of 11.9%.
  
### For K = 20
```{r 11-g-20NN}
knn.auto20 <- knn(train.x2, test.x2, train.y2, k=20)
table(knn.auto20, Auto_testing$mpg01)
mean(knn.auto20 == Auto_testing$mpg01)
```
The accuracy for $K = 20$ is 85.7% or an error rate of 14.3%.  

## Problem 13

**Using the [Boston](http://lib.stat.cmu.edu/datasets/boston) data set, fit classification models in order to predict whether a given suburb has a crime rate above or below the median. Explore logistic regression, LDA, and KNN models using various subset of the predictors. Describe your findings.**  

```{r 13}
attach(Boston)
```
  
Answer.