---
title: "HW 1"
author: "Austin Pesina"
date: "2/2/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## #2

a) This is a regression problem because all the variables have numerical values. This is also an inference problem because we are looking at the relationship between the variables and the CEO’s salary. n = top 500 U.S. firms; p = 3
b) classification; prediction; n = 20; p = 3
c) regression; prediction; n = 52 p=3



## #5

A flexible approach can give a better fit for non-linear models and reduces the bias. On the other hand, it requires a greater number of parameters, it overfits the model, and there is an increase in variance.

A more flexible approach is preferred when we want a prediction while a less flexible approach is preferred when we want an inference and interpretability.


## #6

A parametric approach reduces the problem of estimating f down to one of estimating a set of parameters because it assumes a form for f. 

A non-parametric approach does not assume a particular form of f and requires a large sample to accurately estimate f.

The advantages of a parametric approach over a non-parametric one is that it simplifies the modeling of f down to fewer parameters so not as many observations are needed. The disadvantages are that there is a potentially inaccurate estimate of f if the form of f is wrong; it could also overfit the observations if a more flexible model is used.


## #8
```{r #8, echo=TRUE}
library(ISLR)
data(College)
college <- read.csv("~/Data_Mining/Data/College.csv", header=T)

head(college[, 1:5])

rownames <- college[,1]
fix(college)
college <- college[,-1]
fix(college)

college$Private<-as.factor(college$Private)

summary(college)

pairs(college[, 1:10])

plot(college$Private, college$Outstate, xlab = "Private University",
     ylab ="Out of State tuition in USD", main = "Outstate Tuition Plot")

Elite <-rep("No", nrow(college))
Elite [college$Top10perc >50] = "Yes"
Elite <-as.factor(Elite)
college <- data.frame(college, Elite)
fix(college)
summary(college$Elite)

plot(college$Elite, college$Outstate, xlab = "Elite University",
     ylab ="Out of State tuition in USD", main = "Outstate Tuition Plot")

par(mfrow = c(2,2))
hist(college$Books, col = 2, xlab = "Books", ylab = "Count")
hist(college$PhD, col = 3, xlab = "PhD", ylab = "Count")
hist(college$Grad.Rate, col = 4, xlab = "Grad Rate", ylab = "Count")
hist(college$perc.alumni, col = 6, xlab = "% alumni", ylab = "Count")

summary(college$PhD)

```

## #9
```{r #9, echo=TRUE}
library(ISLR)
data(Auto)
auto <-read.csv("~/Data_Mining/Data/Auto.csv", header=T, na.strings="?")
fix(auto)

str(auto)

quant <- sapply(Auto, is.numeric)
quant
sapply(auto[quant], range)

sapply(auto[quant], mean)
sapply(auto[quant], sd)

sub <-auto[-c(10:85), -c(4,9)]
sapply(sub, range)
sapply(sub, mean)
sapply(sub, sd)

auto$cylinders <-as.factor(auto$cylinders)
auto$year <-as.factor(auto$year)
auto$origin <-as.factor(auto$origin)
pairs(auto[-9])


```
a) All variables except "name" are quantitative.



