---
title: "HW 2"
author: "Austin Pesina"
date: "2/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=F)
```

## Problem 10.
This question should be answered using the [Carseats](https://rdrr.io/cran/ISLR/man/Carseats.html) data set.

```{r getdata}
library(ISLR)
attach(Carseats)
```
**(a) Fit a multiple regression model to predict Sales using Price,Urban,and US.**  
```{r}
fit<-lm(Sales~Price+Urban+US)
summary(fit)
```

**(b) Provide an interpretation of each coefficient in the model. Becareful—some of the variables in the model are qualitative!**  
From the table above, `Price` and `US` are significant predictors of `Sales,` for every \$1 increase in my price, sales decrease by \$54. Sales inside the US are \$1,200 higher than sales outside of the US. Urban has no effect on Sales.

**(c) Write out the model in equation form, being careful to handlethe qualitative variables properly.**  $Sales=13.043469  - 0.054459Price -  0.021916Urban_{Yes} +  1.200573US_{YES}     $

**(d) For which of the predictors can you reject the null hypothesis $H0:β_{j}=0$?**  
`Price` and `US`

**(e) On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there isevidence of association with the outcome.**  
```{r}
fit<-lm(Sales~Price+Urban+US)
summary(fit)
```

**(f) How well do the models in (a) and (e) fit the data?**  
Models do not fit well. Each model explains around 23% of the variance in Sales.

**(g) Using the model from (e), obtain 95 % confidence intervals for the coefficient(s).**  
```{r}
confint(fit)
```
**(h) Is there evidence of outliers or high leverage observations in the model from (e)?**  